Response to the defense feedback

### 2021-06-16 Chapter-1/Chapter-1.tex
Updates from Overleaf

A large portion of side-channel attacks is the joint result of the underlying computer architecture and the software stack.  One example is cache channel attacks. The CPU cache is a smaller but faster memory designed to bridge the gap of the access time between the main memory and CPU registers by storing copies of recently accessed data. Because the size of the cache is significantly smaller than the size of the main memory, the CPU may only find part of the data in the cache. Accessing data from the cache is faster than accessing data from the main memory. Such differences can be reflected in the variance in the program execution time. Hence we have cache channel attacks.

Various countermeasures have been proposed to defend against side-channel attacks. Existing countermeasures can be divided into two categories, hardware-based methods and software-based methods. Hardware-level solutions, such as reducing shared resources, oblivious RAM, and transnational memory~\cite{203878,217537,shih2017t,Zhang:2015:HDL:2775054.2694372}, need to adopt new hardware features and change  today's modern complex computer systems, which are hard to adopt in reality. Therefore, a promising and universal direction is software countermeasures, detecting and eliminating side-channel vulnerabilities in existing software code base. For example, when we examine recent side-channel attacks, many of them exploit two kinds of code patterns: \emph{secret-dependent control-flow transfers} and \emph{secret-dependent data accesses}.
\begin{figure}[h]
    \begin{lstlisting}[xleftmargin=.32\textwidth, xrightmargin=.32\textwidth]
int secret[32];
...
while(i>0){
    r = (r * r) % n;
    if(secret[--i] == 1)
        r = (r * x) % n;   
}
\end{lstlisting}
    \caption{Secret-dependent control-flow transfer}
    \label{fig:secret:cf}
\end{figure}

@@

\begin{figure}[h]
    \begin{lstlisting}[xleftmargin=.32\textwidth, xrightmargin=.32\textwidth]
static char FSb[256] = {...}
... 
uint32_t a = *RK++ ^ \ 
(FSb[(secret)) ^
(FSb[(secret >> 8)] << 8 ) ^
(FSb[(secret >>16)] << 16 ) ^
(FSb[(secret >>24)] << 24 );
...
\end{lstlisting}
    \caption{Secret-dependent data access}
    \label{fig:secret:da}
\end{figure}

@@

However, eliminating side-channel vulnerabilities from real-world software code base is not easy. First, side-channels are a low-level problem in computer systems. As a non-functional property, it can be difficult for programmers to reason about or test.  While recent work~\cite{203878,217537,Wichelmann:2018:MFF:3274694.3274741,Brotzman19Casym,236338,182946} can detect many side-channel vulnerabilities, it either suffers from expensive computation costs or high false positives and false negatives.  Second, most vulnerabilities pose a negligible risk. Although some vulnerabilities result in the key being entirely compromised~\cite{184415, aumuller2002fault}, many others are less severe in reality. Therefore, developers are often reluctant to fixing all of them. We need a proper quantification metric to assess the sensitivity of side-channel vulnerabilities, so a developer can efficiently triage them.  Third, many side-channel vulnerable implementations (e.g., the CRT optimization and lookup tables) have better performance. This is a dilemma between the performance and the security.

@@

\section{Dissertation Organization}
The rest of the dissertation is organized as follows. We present the related work of the dissertation in Chapter~\ref{chapter2}. Chapter~\ref{chapter3} presents a novel method to detect side-channel vulnerabilities in binary executables. Chapter~\ref{chapter4} presents a method that can quantify each side-channel leakage in a single trace attack. Chapter~\ref{chapter5} extends the approach and quantifies the side-channel leakage in a multiple-trace attack.
Chapter~\ref{chapter6} and Chapter~\ref{chapter7} discuss limitations of the dissertation research and propose the future work.

>>>>

A large portion of side-channel attacks is the joint result of the underlying computer architecture and the software stack.  One example is cache channel attacks. The CPU cache is a smaller but faster memory to bridge the gap of the access time between the main memory and CPU registers by storing copies of recently accessed data. Because the size of the cache is significantly smaller than the size of the main memory, the CPU may only find part of the data in the cache. Accessing data from the cache is faster than accessing data from the main memory. Such differences can be reflected in the variance in the program execution time. Hence we have cache channel attacks.

Various countermeasures have been proposed to defend against side-channel attacks. Existing countermeasures can be divided into two categories, hardware-based methods and software-based methods. Hardware-level solutions, such as reducing shared resources, oblivious RAM, and transnational memory~\cite{203878,217537,shih2017t,Zhang:2015:HDL:2775054.2694372}, need to adopt new hardware features and change  today's modern complex computer systems, which are hard to adopt in reality. Therefore, a promising and universal direction is software countermeasures, detecting and eliminating side-channel vulnerabilities in existing software codebase. For example, when we examine recent side-channel attacks, many of them exploit two kinds of code patterns: \emph{secret-dependent control-flow transfers} and \emph{secret-dependent data accesses}.
\begin{figure}[h]
    \begin{lstlisting}[xleftmargin=.3\textwidth, xrightmargin=.32\textwidth]
int secret[32];
...
while(i>0){
    r = (r * r) % n;
    if(secret[--i] == 1)
        r = (r * x) % n;   
}
\end{lstlisting}
    \caption{Secret-dependent control-flow transfer}
    \label{fig:secret:cf}
\end{figure}

@@

\begin{figure}[h]
    \begin{lstlisting}[xleftmargin=.3\textwidth, xrightmargin=.2\textwidth]
static char FSb[256] = {...}
... 
uint32_t a = *RK++ ^ \ 
(FSb[(secret)) ^
(FSb[(secret >> 8)] << 8 ) ^
(FSb[(secret >>16)] << 16 ) ^
(FSb[(secret >>24)] << 24 );
...
\end{lstlisting}
    \caption{Secret-dependent data access}
    \label{fig:secret:da}
\end{figure}

@@

However, eliminating side-channel vulnerabilities from real-world software codebase is not easy. First, side-channels are a low-level problem in computer systems. As a non-functional property, it can be difficult for programmers to reason about or test.  While recent work~\cite{203878,217537,Wichelmann:2018:MFF:3274694.3274741,Brotzman19Casym,236338,182946} can detect many side-channel vulnerabilities, it suffers from either expensive computation costs or high false positives and false negatives.  Second, most vulnerabilities pose a negligible risk. Although some vulnerabilities result in the key being entirely compromised~\cite{184415, aumuller2002fault}, many others are less severe in reality. Therefore, developers are often reluctant to fixing all of them. We need a proper quantification metric to assess the sensitivity of side-channel vulnerabilities, so a developer can efficiently triage them.  Third, many side-channel vulnerable implementations (e.g., using the Chinese Remainder Theorem optimization in RSA and lookup tables in AES) have better performance. This is a dilemma between the performance and the security.

@@

\section{Dissertation Organization}
The rest of the dissertation is organized as follows. We present the related work of the dissertation in Chapter~\ref{chapter2}. Chapter~\ref{chapter3} presents a novel method to detect side-channel vulnerabilities in binary executables. Chapter~\ref{chapter4} presents a method that can quantify each side-channel leakage in a single trace attack. Chapter~\ref{chapter5} extends the approach and quantifies the side-channel leakage in a multiple-trace attack.
Chapter~\ref{chapter6} and Chapter~\ref{chapter7} discuss limitations and future work.

### 2021-06-16 SupplementaryMaterial/Abstract.tex
Updates from Overleaf

\vspace{-0.3in}
Side channels are ubiquitous in modern computer systems as sensitive
information can leak through many mechanisms such as power consumption, 
execution timing, and even electromagnetic radiation. Among them, 
address-based side-channel attacks, such as cache-based attacks, 
memory page attacks, and controlled-channel attacks, are especially
problematic as they do not require physical proximity. Hardware 
countermeasures, which usually require changes to the complex 
underlying hardware, are hard to adopt in practice.  On the contrary, 
software approaches are generally easy to implement. While some existing 
tools can detect side-channel leakages, many of these approaches are 
computationally expensive or imprecise. Besides, many such vulnerabilities 
leak a negligible amount of sensitive information, and thus developers 
are often reluctant to address them. Existing tools do not provide sufficient 
information, such as the amount of information leakage through side channels, 
to evaluate the severity of a vulnerability.

@@

We implement the approaches and apply them to popular cryptography libraries.  
The evaluation results confirm that our side-channel detection method 
is much faster than state-of-art tools while identifying all the known leakages 
reported by previous tools.
The experiments also show that 
our side-channel analysis reports quite precise leakage information
that can help developers better triage the reported vulnerabilities.
This dissertation research develops fundamental and practical techniques for precise side-channel 
analysis in software systems. We have also open sourced our research software prototypes.
As a result, developers can use our tools to develop more secure systems and the academic and industry
communities can further advance side-channel analysis on top of our research.

>>>>

\vspace{-0.3in}
Side channels are ubiquitous in modern computer systems as sensitive
information can leak through many mechanisms such as power consumption, 
execution time, and even electromagnetic radiation. Among them, 
address-based side-channel attacks, such as cache-based attacks, 
memory page attacks, and controlled-channel attacks, are especially
problematic as they do not require physical proximity. Hardware 
countermeasures, which usually require changes to the complex 
underlying hardware, are hard to adopt in practice.  On the contrary, 
software approaches are generally easy to implement. While some existing 
tools can detect side-channel leakages, many of these approaches are 
computationally expensive or imprecise. Besides, many such vulnerabilities 
leak a negligible amount of sensitive information, and thus developers 
are often reluctant to address them. Existing tools do not provide sufficient 
information, such as the amount of information leaked through side channels, 
to evaluate the severity of a vulnerability.

@@

We implement the approaches and apply them to popular cryptography libraries.  
The evaluation results confirm that our side-channel detection method 
is much faster than state-of-art tools while identifying all the known leakages 
reported by previous tools.
The experiments also show that 
our side-channel analysis reports quite precise leakage information
that can help developers better triage the reported vulnerabilities.
This dissertation research develops fundamental and practical techniques for precise side-channel 
analysis in software systems. We have also released our research software prototypes.
As a result, developers can use our tools to develop more secure systems and the academic and industry
communities can further advance side-channel analysis on top of our research.
